A Parallel General Purpose Multi-Objective Optimization Framework, with Application To Beam Dynamics

### 2018-07-10 paper/theor-experiment.tex
more edits to awa paper section

A study on time step and 
number of particles was done to reduce the time to simulation while 
maintaining the physics of interest. 
The grid size $16 \times 16 \times 32$ was chosen, 
and parallelized in the x and y directions.
After comparing several options we choose ten thousand particles 
in combination with adjusting the time step (dT) near sensitive elements 
(i.e. quadurpoles, kicker). 
The resulting simulations are low fidelity, but closely approximate 
the mid fidelity simulations for metrics of interest, as shown in 
Fig. \ref{tstep}. For all models, the longitudinal parameters (rms$_s$ and energy) 
are calculated correctly, but we see discrepancies in the transverse 
rms$_x$ and $\epsilon_x$ in the low fidelity results. The
average run time of each simulation with the adjusted time steps was 1.6~minutes.
In comparison the mid fidelity simulations ran for 18~minutes.
%A summary of the parameters used is listed in Table \ref{fidelity}.

@@

\subsubsection{AWA Optimization Results}
While the optimization problem and goals were well defined (Section \ref{awaproblem}), 
it was not clear what the best input parameters for the NSGA-II should be.
These parameters include gene mutation probability, mutation probability, 
recombination probability, number of individuals (initial \lsnote{you have an extra `and' here} and for later generations), 
and number of generations to complete. 
Given the beam line and design variables given in Section \ref{awaproblem}, 
four optimization experiments were done with varying NSGA-II parameters. 
Similar to the time step scan, 
the goal of this exercise was to determine which set of optimization
parameters strongly influence the results, 
and whether there was a time to solution difference.
From here on, we will reference each experiment as ex-1, ex-2, ex-3, and ex-4
as shown in Table \ref{extable}. 
\begin{table}%[h!]
	\begin{center}
		\caption{Input Parameters for initial twenty four hour AWA optimization experiments. 
			The gene mutation probability was equal to the mutation probability (not shown) in all four experiments. 
			The max number of individuals per generation was~80.}
		\label{extable}
		\begin{tabular*}{\textwidth}{l @{\extracolsep{\fill}} C C D }
			\hline\noalign{\smallskip}
			& Gene Mutation Probability & Recombination Probibility & Number of completed generations \\
			\noalign{\smallskip}\hline\noalign{\smallskip}
			ex-1 &  0.1  & 0.9  &  96 \\
			ex-2 &  0.3  & 0.7  &  81 \\
			ex-3 &  0.8  & 0.2  &  53 \\
			ex-4 &  0.01 & 0.09 &  95 \\
			\noalign{\smallskip}\hline
		\end{tabular*}
	\end{center}
\end{table}

The maximum number of individuals per generation was fixed at 80, 
Each experiment was allowed to run for twenty four hours, with 
a maximum generation limit of 100. 
We reduced the seven objectives to the four needed to construct 
the Pareto fronts that would be the final goal of this study.
The objectives include: $\varepsilon_{x}\left(s = s_1\right)\text{, } \varepsilon_{x}\left(s = s_2\right)\text{, } \text{rms}_{s}\left(s = s_1\right)\text{, and }  \text{rms}_{s}\left(s = s_2\right)$. 
\lsnote{You could define $\varepsilon_{x1}$ to be $\varepsilon_{x} (s=s1)$ and similarly in the table and then use going forward to simplify notation}
The optimization parameters specific to OPAL input for ex-3 \lsnote{are} given as an example.
Note the design variables defined in Section \ref{awaproblem}
precede the optimization parameters in the OPAL input file.

@@

@@

Next compare the design variables at each point on the 
Pareto front.
\begin{figure}
	\centering
	\begin{subfigure}{1\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Report/ex-pareto1}
		\caption{Pareto fronts at $s_1$.}
		\label{bounds1}
	\end{subfigure}
	\begin{subfigure}{1\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Report/ex-pareto2}
		\caption{Pareto fronts at $s_2$.}
		\label{bounds2}
	\end{subfigure}
	\caption{Comparison of design variables for each Pareto front.}
	\label{bounds}
\end{figure}

The initial number of individuals was fixed at 656, 
and the minimum number individuals in later generations was fixed at 328. 
These numbers \lsnote{were} in part based on the architecture of the KNL's. 
Since each simulation takes 8 cores, and there are 64 cores per KNL node, 
we wanted a large population size that would fit evenly on these resources.

>>>>

\subsubsection{Preliminary Tests} \label{awa:subsection:test}
\nrnote{What else could we name this subsection?}
A study on time step and 
number of particles was done to reduce the time to simulation while 
maintaining the physics of interest. 
The grid size $16 \times 16 \times 32$ was chosen, 
and parallelized in the x and y directions.
After comparing several options we choose ten thousand particles 
in combination with adjusting the time step (dT) near sensitive elements 
(i.e. quadurpoles, kicker). 
The resulting simulations are low fidelity, but closely approximate 
the mid fidelity simulations for metrics of interest, as shown in 
Fig. \ref{tstep}. For all models, the longitudinal parameters (rms$_s$ and energy) 
are calculated correctly, but we see discrepancies in the transverse 
rms$_x$ and $\epsilon_x$ in the low fidelity results. The
average run time of each simulation with the adjusted time steps was 1.6~minutes.
In comparison the mid fidelity simulations ran for 18~minutes.
%A summary of the parameters used is listed in Table \ref{fidelity}.

@@

While the optimization problem and goals were well defined (Section \ref{awaproblem}), 
it was not clear what the best input parameters for the NSGA-II should be.
These parameters include gene mutation probability, mutation probability, 
recombination probability, number of individuals, 
and number of generations to complete. 
Given the beam line and design variables given in Section \ref{awaproblem},
\nrnote{Changed variables for final run, the variables in 5.3 no longer match these small runs. Need to rework this section} 
four optimization experiments were done with varying NSGA-II parameters. 
Similar to the time step scan, 
the goal of this exercise was to determine which set of optimization
parameters strongly influence the results, 
and whether there was a time to solution difference.
From here on, we will reference each experiment as ex-1, ex-2, ex-3, and ex-4
as shown in Table \ref{extable}. 
\begin{table}%[h!]
	\begin{center}
		\caption{Input Parameters for initial twenty four hour AWA optimization experiments. 
			The gene mutation probability was equal to the mutation probability (not shown) in all four experiments. 
			The max number of individuals per generation was~80.}
		\label{extable}
		\begin{tabular*}{\textwidth}{l @{\extracolsep{\fill}} C C D }
			\hline\noalign{\smallskip}
			& Gene Mutation Probability & Recombination Probibility & Number of completed generations \\
			\noalign{\smallskip}\hline\noalign{\smallskip}
			ex-1 &  0.1  & 0.9  &  96 \\
			ex-2 &  0.3  & 0.7  &  81 \\
			ex-3 &  0.8  & 0.2  &  53 \\
			ex-4 &  0.01 & 0.09 &  95 \\
			\noalign{\smallskip}\hline
		\end{tabular*}
	\end{center}
\end{table}

The maximum number of individuals per generation was fixed at 80, 
Each experiment was allowed to run for twenty four hours, with 
a maximum generation limit of 100. 
We reduced the seven objectives to the four needed to construct 
the Pareto fronts that would be the final goal of this study.
The objectives include: $\varepsilon_{x}\left(s = s_1\right)\text{, } \varepsilon_{x}\left(s = s_2\right)\text{, } \text{rms}_{s}\left(s = s_1\right)\text{, and }  \text{rms}_{s}\left(s = s_2\right)$. 
\lsnote{You could define $\varepsilon_{x1}$ to be $\varepsilon_{x} (s=s1)$ and similarly in the table and then use going forward to simplify notation}
The optimization parameters specific to OPAL input for ex-3 are given as an example.
Note the design variables defined in Section \ref{awaproblem}
precede the optimization parameters in the OPAL input file.

@@

Next compare the design variables at each point on the 
Pareto front.

@@

The initial number of individuals was fixed at 656, 
and the minimum number individuals in later generations was fixed at 328. 
These numbers were in part based on the architecture of the KNL's. 
Since each simulation takes 8 cores, and there are 64 cores per KNL node, 
we wanted a large population size that would fit evenly on these resources. 
This would be the population sizes used for all subsequent runs.

\subsubsection{AWA Optimization Results}
With the simulation, GA, and node resource parameters set by the work in \ref{awa:subsection:test}, 
the full optimization problem described in \ref{awaproblem} was run for 200 generations.

\begin{figure}
	%\includegraphics[width=0.5\textwidth]{}\includegraphics[width=0.5\textwidth]{}
	\label{beamsizes}
\end{figure}
